Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-10 popl24/bruteforce.tex
checkpoint

@@

As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
$\runs(\prog)$ is exponential in the amount of randomness used in the
protocol which generally can be used as a proxy for the protocol
size. In Section \ref{section-composition} we will show how the
methods described here can be used to individually verify
compositional properties of small program components, that guarantee
security properties in larger protocols that combine them. Thus, while
scalability to program size is a clear challenge for our brute force
techniques alone, our ultimate goal is to use them for verification of
compositional properties in larger programs.

>>>>

\begin{fpfig}[t]{Truth table visualizations of $\runs(\prog_{\ref{example-otp}})$ (L) and
    $\runs(\prog_{\ref{example-lambda-obliv}(a)})$ (R).}{fig-basic-distributions}
{\footnotesize
  $$
  \begin{array}{cc}
    \begin{array}{cccccc}
      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
      \hline
      0 & 0 & 0 & 0 & 0 & 0 \\ 
      0 & 1 & 1 & 1 & 0 & 1 \\ 
      1 & 0 & 0 & 1 & 1 & 1 \\ 
      1 & 1 & 1 & 0 & 1 & 0
    \end{array}
    & 
    \begin{array}{ccccc}
      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
      \hline
      0 & 0 & 0 & 0 & 0 \\ 
      0 & 0 & 1 & 1 & 0 \\ 
      0 & 1 & 0 & 0 & 1 \\ 
      0 & 1 & 1 & 1 & 1 \\
      1 & 0 & 0 & 0 & 0 \\ 
      1 & 0 & 1 & 0 & 0 \\ 
      1 & 1 & 0 & 1 & 1 \\ 
      1 & 1 & 1 & 1 & 1  
    \end{array}
  \end{array}
  $$
}
\end{fpfig}

@@

As we've observed, $\runs(\prog)$ can be represented in a truth table
form, as illustrated in Figure \ref{fig-basic-distributions}.

### 2023-07-10 popl24/pmf.tex
checkpoint

We consider random tapes to be selected from a uniform distribution.
In our threat model, we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this, we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pmf(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pmf}{S\cup V}
  $$
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

@@

\begin{fpfig}[t]{Truth table visualization of basic distributions for Examples \ref{example-otp} (L) and \ref{example-lambda-obliv}(a) (R).}{fig-basic-distributions}
{\footnotesize
  $$
  \begin{array}{cc}
    \begin{array}{cccccc}
      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
      \hline
      0 & 0 & 0 & 0 & 0 & 0 \\ 
      0 & 1 & 1 & 1 & 0 & 1 \\ 
      1 & 0 & 0 & 1 & 1 & 1 \\ 
      1 & 1 & 1 & 0 & 1 & 0
    \end{array}
    & 
    \begin{array}{ccccc}
      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
      \hline
      0 & 0 & 0 & 0 & 0 \\ 
      0 & 0 & 1 & 1 & 0 \\ 
      0 & 1 & 0 & 0 & 1 \\ 
      0 & 1 & 1 & 1 & 1 \\
      1 & 0 & 0 & 0 & 0 \\ 
      1 & 0 & 1 & 0 & 0 \\ 
      1 & 1 & 0 & 1 & 1 \\ 
      1 & 1 & 1 & 1 & 1  
    \end{array}
  \end{array}
  $$
}
\end{fpfig}

\paragraph{Example \ref{example-otp} revisited} For small programs it is
helpful to visualize their basic distributions graphically as their
truth tables, where each row corresponds to an equally likely final
memory for the given program. In Figure \ref{fig-basic-distributions}
we show this for Example \ref{example-otp}. In this truth table we can
see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated. But
from the adversarial view, $\sx{1}{0}$ is 1 or 0 with equal
probability given any value of $\vx{0}{0}$. More precisely, for any
$\beta_1,\beta_2$ we have:
\begin{mathpar}
(\progd(\prog_{\ref{example-otp}}))(\{\sx{1}{0} \mapsto \beta_1 \}|\{ \vx{0}{0} \mapsto \beta_2 \}) = .5
\end{mathpar}
Note that this corresponds to \emph{perfect secrecy} as described
in \cite{barthe2019probabilistic}.

\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
to leak information probabilistically through observable dependencies
in views. In the basic distribution for Example \ref{example-lambda-obliv}(a)
illustrated in Figure \ref{fig-basic-distributions}, we can see that
the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
three different runs of the protocol, and in 2 out of 3 the
honest secret $\sx{1}{0}$ is 1. More precisely:
$$
(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
$$
An important point illustrated by this example is that the
probability of values in views, and not just the values themselves,
can release information to the adversary.

>>>>

We consider random tapes to be selected from a uniform distribution.
In our threat model, we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this, we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pmf(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pmf}{S\cup V}
  $$
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

@@

\paragraph{Example \ref{example-otp} revisited}
%For small programs it is
%helpful to visualize their basic distributions graphically as their
%truth tables, where each row corresponds to an equally likely final
%memory for the given program. In Figure \ref{fig-basic-distributions}
%we show this for Example \ref{example-otp}. In this truth table we can
%see that $\sx{1}{0}$ and $\vx{2}{2}$ are perfectly correlated.
In this example, 
from the adversarial view, $\sx{1}{0}$ is 1 or 0 with equal
probability given any value of $\vx{0}{0}$. More precisely, for any
$\beta_1,\beta_2$ we have:
\begin{mathpar}
(\progd(\prog_{\ref{example-otp}}))(\{\sx{1}{0} \mapsto \beta_1 \}|\{ \vx{0}{0} \mapsto \beta_2 \}) = .5
\end{mathpar}
Note that this corresponds to \emph{perfect secrecy} as described
in \cite{barthe2019probabilistic}.

\paragraph{Example \ref{example-lambda-obliv}(a) revisited} As demonstrated
by Example \ref{example-lambda-obliv} taken from \cite{darais2019language}, it is possible
to leak information probabilistically through observable dependencies
in views.
%In the basic distribution for Example \ref{example-lambda-obliv}(a)
%illustrated in Figure \ref{fig-basic-distributions},
We can see that
the adversarial views $\vx{0}{0}$ and $\vx{0}{1}$ are both 1 in
three different runs of the protocol, and in 2 out of 3 the
honest secret $\sx{1}{0}$ is 1. More precisely:
$$
(\progd(\prog_{\ref{example-lambda-obliv}(a)}))(\{\sx{1}{0} \mapsto 1 \}|\{ \vx{0}{0} \mapsto 1, \vx{0}{1} \mapsto 1 \}) = 2/3
$$
An important point illustrated by this example is that the
probability of values in views, and not just the values themselves,
can release information to the adversary.
